---
title: "Monte carlo"
author: "Dan Weinberger"
date: "February 28, 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rsatscan)
library(dplyr)
```

## The problem
We have 2 regions, each with 25 municipalities each. We measure the number of cases per municipality and want to see if the average number of cases differs by region

Lets start by simulating some data. Region 1 has an average of 10 cases/muni; region 2 has 13 cases per muni
```{r cars}
set.seed(123)
region1<-rpois(n=25, lambda=10)

region2<-rpois(n=25, lambda=13)
```

```{r}
trans.red<-rgb(1,0,0, alpha=0.5)
trans.blue<-rgb(0,0,1, alpha=0.5)

hist(region1, col=trans.red, main='Histogram in the 2 regions', xlim=c(0,20), ylim=c(0,10))
hist(region2, add=T, col=trans.blue)
```

#Calculate the mean number of cases in each region and their difference
```{r}
mean(region1)
mean(region2)

observed.diff <- mean(region2)-mean(region1)
observed.diff
```

## Combine data from the two regions into a data frame

```{r}
region1.df<-cbind.data.frame('counts'=region1,'region'=rep(1, length(region1)))

region2.df<-cbind.data.frame('counts'=region2,'region'=rep(2, length(region2)))

combined.df<-rbind.data.frame(region1.df,region2.df)
```

#Now again calculate the summary statistics

```{r}
diff.mean.true <- combined.df %>%
  group_by(region) %>%
  summarize(ave_counts =mean(counts))

diff.mean.true$ave_counts[2]- diff.mean.true$ave_counts[1]

```

#Now scramble the region assignment column
Let's rerun it a few times and see what happens
```{r}
combined.df.shuffle<-combined.df

combined.df.shuffle$reorder<- sample(combined.df.shuffle$region) 

#create a reshuffled version of region assigment

diff.mean.shuffle <- combined.df.shuffle %>%
  group_by(reorder) %>%
  summarize(ave_counts =mean(counts))

diff.mean.shuffle$ave_counts[2]- diff.mean.shuffle$ave_counts[1]

```

#Now let's do it 999 times
Easy way to do this is to make a function, and call this 999 times

```{r}
mc_function <- function(ds){
  combined.df.shuffle <- ds
  
  combined.df.shuffle$reorder<- sample(combined.df.shuffle$region) 
  
  #create a reshuffled version of region assigment
  
  diff.mean.shuffle <- combined.df.shuffle %>%
    group_by(reorder) %>%
    summarize(ave_counts =mean(counts))
  
  test.stat <-diff.mean.shuffle$ave_counts[2]- diff.mean.shuffle$ave_counts[1]
  
  return(test.stat)
}

```

Call the function 999 times

```{r}
set.seed(123)
shuffled.diff <- replicate(999, mc_function(combined.df))
```

```{r}
hist(shuffled.diff)
abline( v=observed.diff, col='red', lty=2)
```

## Calculating a P-value
Determine what percent of the simulations had a value that was larger than the observed values
```{r}
N.greater.obs <- sum(shuffled.diff > observed.diff ) #How many of simulations were greater than observed?

p.value <- N.greater.obs/ length(shuffled.diff) #What proportion of simulations greater than observed?

print('P-value')

p.value
```


## What does SATSCAN do?
For the space-time permutation,
1) Calculating the scan statistic over all time window and geographic region combinations
2) Scramble the dates and calculate the statistics on the scrambled datasets
3) Repeat 999 times
4) compare the most extreme scan statistic from step 1 with the distribution from steps 2-3 

```{r}
head(NYCfevercas)

scramble1<-NYCfevercas
scramble1$scampledate<-sample(scramble1$date)
head(scramble1)
```

