---
title: "Monte carlo"
author: "Dan Weinberger"
date: "February 19, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(rsatscan)
```

## The problem
We have 2 regions, each with 25 municipalities each. We measure the number of cases per municipality and want to see if the average number of cases differs by region

Lets start by simulating some data. Region 1 has an average of 10 cases/muni; region 2 has 12 cases per muni
```{r cars}
region1<-rpois(n=25, lambda=10)
region2<-rpois(n=25, lambda=12)

trans.red<-rgb(1,0,0, alpha=0.5)
trans.blue<-rgb(0,0,1, alpha=0.5)

hist(region1, col=trans.red, main='Histogram in the 2 regions')
hist(region2, add=T, col=trans.blue)
```

#Calculate the mean number of cases in each region and their difference
```{r}
mean(region1)
mean(region2)

mean(region2)-mean(region1)

```

#Combine data from the two regions into a data frame
```{r}
region1.df<-cbind.data.frame('counts'=region1,'region'=rep(1, length(region1)))
region2.df<-cbind.data.frame('counts'=region2,'region'=rep(2, length(region2)))
combined.df<-rbind.data.frame(region1.df,region2.df)
```

#Now again calculate the summary statistics
This time we will use split, apply, combine approach on the data frame
```{r}
df.split<-split(combined.df, combined.df$region)
region.mean<-sapply(df.split, function(x) mean(x$counts))

diff.mean.true<-region.mean[2]- region.mean[1]

```

#Now scramble the region assignment column
Let's rerun it a few times and see what happens
```{r}
combined.df.shuffle<-combined.df
combined.df.shuffle$reorder<- sample(combined.df.shuffle$region) #create a reshuffled version of region assigment
#combined.df.shuffle$reorder
df.split.shuffle<-split(combined.df.shuffle, combined.df.shuffle$reorder)
region.mean<-sapply(df.split.shuffle, function(x) mean(x$counts))
diff.mean<-region.mean[2]- region.mean[1]
diff.mean
```

#Now let's do it 999 times
```{r}
set.seed(123)
diff.mean.mc<-rep(NA, 999) #Make a vector of missing values
for(i in 1:999){
combined.df.shuffle<-combined.df
combined.df.shuffle$reorder<- sample(combined.df.shuffle$region) #create a reshuffled version of region assigment
#combined.df.shuffle$reorder
df.split.shuffle<-split(combined.df.shuffle, combined.df.shuffle$reorder)
region.mean<-sapply(df.split.shuffle, function(x) mean(x$counts))
diff.mean.mc[i]<-region.mean[2]- region.mean[1]
}
hist(diff.mean.mc ,xlim=c(-2, 5))
abline(v=diff.mean.true, col='red', lty=2)
```

## Calculating a P-value
Determine what percent of the simulations had a value that was larger than the observed values
```{r}
N.greater.obs<-sum(diff.mean.mc > diff.mean.true ) #How many of simulations were greater than observed?
p.value<- N.greater.obs/ length(diff.mean.mc) #What proportion of simulations greater than observed?
print('P-value')
p.value
```


## What does SATSCAN do?
For the space-time permutation,
1) Calculating the scan statistic over all time window and geographic region combinations
2) Scramble the dates and calculate the statistics on the scrambled datasets
3) Repeat 999 times
4) compare the most extreme scan statistic from step 1 with the distribution from steps 2-3 

```{r}
head(NYCfevercas)

scramble1<-NYCfevercas
scramble1$scampledate<-sample(scramble1$date)
head(scramble1)
```

