---
title: "rsatscan intro"
author: "Dan Weinberger"
date: "February 19, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath(".."))
#install.packages('rsatscan')
#install.packages('rgdal')
library(rsatscan)
library(rgdal)

```

## Data
The NYC fever data come pre-packaged with rsatscan. See the files below

```{r cars}
head(NYCfevergeo)
head(NYCfevercas)
```

Read in shape file for mapping
"dsn"" gives the folder where the .shp file is saved; layer gives the name of the .shp file. Note the folder must contain the .shp files as well as the .prj, .shx, .dbf files.
```{r}
getwd()
shape.NYC <- readOGR(dsn = "./NYC shape files", layer = "geo_export_ef7f38de-fdc9-46c6-bf6b-36dcb40dfda8")
```

## Specify the parameters for SATSCAN and generate the input files needed
Save in a temporary directory
```{r}
invisible(ss.options(reset=TRUE)) #Reset parameters
```

Parameters for SATSCAN
```{r}
ss.options(list(CaseFile="NYCfever.cas", PrecisionCaseTimes=3))
ss.options(c("StartDate=2001/11/1","EndDate=2001/11/24"))
ss.options(list(CoordinatesFile="NYCfever.geo", AnalysisType=4, ModelType=2, TimeAggregationUnits=3))
ss.options(list(UseDistanceFromCenterOption="y", MaxSpatialSizeInDistanceFromCenter=3, NonCompactnessPenalty=0))
ss.options(list(MaxTemporalSizeInterpretation=1, MaxTemporalSize=7))
ss.options(list(ProspectiveStartDate="2001/11/24", ReportGiniClusters="n", LogRunToHistoryFile="n"))
```

Where to savethe output?
```{r}
td = tempdir()
write.ss.prm(td, "NYCfever")
write.cas(NYCfevercas, td, "NYCfever")
write.geo(NYCfevergeo, td, "NYCfever")
```


#Run SATSCAN
Save results in an object called "NYCfever"
Need to point to directory on your computer where SATSCAN program is saved
```{r}
NYCfever = satscan(td, "NYCfever", sslocation="C:/Program Files (x86)/SaTScan")

```

## Summary of results

```{r}
summary(NYCfever)
```

#Cluster info
```{r}
#See the main results and info about each cluster
NYCfever$main
```

#Map the results
```{r}
sp::plot(shape.NYC)  #Map NYC borough boundaries
sp::plot(NYCfever$shapeclust, add=T, col='red', pch=16) #Overlay the SATSCAN cluster results
```
## View the SATSCAN results compared to Monte Carlo resampling
```{r}
#View the clusters vs the Monte Carlo simulations to see how unlikely the result is
hist(unlist(NYCfever$llr), main="Monte Carlo")
abline(v=NYCfever$col[,c("TEST_STAT")], col = "red")
```
